The Inner Mass-Ratio Distribution of Intermediate-Mass Stars

### 2016-03-08 paper/ms.tex
Updated abstract

\begin{abstract}
Binary stars and higher-order multiple systems are a ubiquitous outcome of star formation, especially as the system mass increases. The distribution of mass ratios is a unique probe into the conditions of the collapsing cloud core and circumstellar disk(s) of the binary fragments. Inside $a \sim 1000$ AU the disks from the two forming stars can interact, and additionally companions can form directly through disk fragmentation. We should therefore expect the mass-ratio distribution of close companions to differ from that of wide companions. This prediction is difficult to test with intermediate-mass to massive primary stars because close companions are so much dimmer. We present the results of a survey searching for companions to A- and B-type stars using the direct spectral detection technique, which is sensitive to late-type companions within $\sim 1''$ of the primary and which has no inner working angle. We estimate the temperatures and masses of the 57 companions , 25 of which are new detections. We find that the mass-ratio distribution for our sample has a robust maximum near $q \sim 0.3$. Our mass-ratio distribution has a very different form than in previous work, where it is usually well-described by a power law, and indicates that close companions to intermediate-mass stars experience significantly different accretion histories or formation mechanisms than wide companions.
\end{abstract}
\keywords{binaries: spectroscopic, stars:early-type, stars:formation, stars:statistics}

@@

The overall multiplicity rate and the distributions of mass ratio, period, and eccentricity of a binary star population place important constraints on the mode of binary star formation. While the period and eccentricity are altered by dynamical processing in the birth cluster, the present-day mass ratio of a binary system is a direct result of its formation \citep{Parker2013}. Most binary stars are thought to form via core fragmentation \citep{Boss1979, Boss1986, Bate1995}, in which a collapsing core fragments into two or more individual protostars. The number and initial masses of the fragments are set by the total core mass, as well as its rotation, turbulence, and its temperature and density structure. If the fragments are well separated ($a \gtrsim 1000$ AU), they will evolve independently of each other, accreting mass from the core material onto their own protostellar disks and then onto the protostars themselves. However close fragments ($a \sim 100$ AU) will interact with each other; the protostellar disk may be truncated, destabilized, or form into a circumbinary disk if the separation is small enough \citep{Bate1997}. In addition, an unstable disk can fragment to form low-mass companions \citep{Kratter2006, Stamatellos2011}. The mass ratios of close companions formed via either mechanism should be affected by preferential accretion. Most work has suggested that the disk material will preferentially accrete onto the lower mass companion \citep{Bate1997, BBB2002}; however, recent work has indicated that magnetic disk braking may result in preferential accretion onto the more massive component \citep{Zhao2013} instead. In either case, we would expect to find a mass ratio distribution for companions inside a few $100$ AU that differs from that of companions on wider orbits.

The mass ratio, period, and eccentricity distributions are fairly well-known for solar type stars \citep{Duquennoy1991, Raghavan2010} and cooler stars \citep{Fischer1992, Delfosse2004}. Interestingly, the mass ratio distribution appears to be invariant to separation for these stars \citep{Meyer2013}, contrary to the theoretical expectations. All of the distributions are much less certain for more massive stars. The reason for this is two-fold: first, more massive stars tend to more rare and farther away, meaning many of the companions are angularly close to the very bright primaries and difficult to detect with imaging techniques. Second, the primary stars tend to be rapid rotators, which limits radial velocity precision to $\sim 1 \mathrm{km\ s}^{-1}$. Additionally, radial velocity monitoring can only measure a mass ratio if spectral lines from both components are visible; this typically suffers from the same flux ratio difficulty as imaging techniques.

Nonetheless, \citet{DeRosa2014} performed an adaptive optics imaging survey of nearby A-type stars, and found that the mass ratio distribution is well-described by a power law with large slope, indicating a very strong preference for low-mass companions. They also found initial evidence that the mass ratio distribution for companions inside $125$ AU has a much shallower power law slope than that of wide companions, and is consistent with flat. Their close companion subsample contained only 18 binary systems, and the result is complicated by the inherent difficulty of detecting close companions with low mass ratios in an imaging survey.

@@

In this paper, we describe a spectroscopic survey of nearby chemically normal, main sequence intermediate-mass stars ($M \approx 1.5 - 15 M_{\odot}$). We search for companions using the direct spectral detection technique \citep{Gullikson2016}, which has a separation-invariant detection rate for all separations inside $\sim 1 ''$. We describe the stellar sample and data used for the survey in Section \ref{sec:obs}, as well as the data reduction steps in the same section. Next, we describe the direct spectral detection method and tabulate the companion detections in Section \ref{sec:companions}. We estimate the mass and age of the primary stars in Section \ref{sec:sp}, and discuss the survey completeness in Section \ref{sec:completeness}. Finally, we end with a derivation of the mass ratio distribution from our sample in Section \ref{sec:mrd} and discuss its implications for binary formation in Section \ref{sec:discussion}.

@@

We use the methodology developed in \citet{Foreman2014} to perform bayesian inference on the shape and form of the mass ratio distribution. The log-likelihood function in this formalism is derived from modeling the survey as a draw from the inhomogeneous Poisson process with rate density $\Gamma \equiv KQ(q)P(q)$:

@@

\begin{align}
P(\vec{\theta}| \alpha, m, \tau, \epsilon) &= \mathcal{N}(\vec{\theta} | m, K(\{\Delta_j\}, \alpha, \tau, \epsilon)) \\
K_{ij} &= \sqrt{\left[\alpha \exp{\left(-\frac{(\Delta_i - \Delta_j)^2}{2\tau^2}\right)}\right]^2 + \epsilon^2 \delta_{ij}}
\end{align}
The smoothing prior is an 7-dimensional gaussian with mean $m$ and covariance matrix $K_{ij}$, and encodes our belief that the mass ratio distribution is a smoothly varying function while leaving enough flexibility to let the data drive the shape of the function. Since we have introduced three new hyperparameters ($a, m, \tau, \epsilon$), we must apply a prior to them and marginalize over them when estimating the bin heights. We choose log-uniform priors for $a, \tau$, and $\epsilon$, and a uniform prior for the mean $m$. The full posterior probability distribution for the histogram model is:

@@

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 300 samples of the lognormal distribution fit to show the spread allowed by the data. The best-fit power law is plotted with a red dot-dashed line. We also estimate the mass ratio distribution expected from random pairing of the Kroupa Initial Mass Function (IMF), and show the result in yellow. We estimated the distribution by drawing 100000 primary stars from the IMF with masses between $1.5 < M/M_{\odot} < 20$. We then drew companions from the same IMF, with the restriction that the companion has a lower mass than the primary. The result plotted in yellow in Figure \ref{fig:mrd} is a gaussian kernel density estimate of the resulting mass ratios, with a bandwidth of $0.05$.

\begin{figure}
\includegraphics[width=\columnwidth]{MRD_total.pdf}
\caption{Mass ratio distribution for our sample. The data was fit to a histogram, a lognormal distribution, and a power law. The histogram is shown in the solid blue blocks, with $1 \sigma$ uncertainties marked with error bars. The variance of the lognormal  fit is shown with 300 samples from the posterior probability distribution for the parameters in green. We also show the best-fit power law and the mass ratio distribution resulting from random pairing of the Kroupa initial mass function. Finally, we display the raw mass ratio measurements with associated uncertainties in the cluster of data points near the top of the figure.}
\label{fig:mrd}
\end{figure}

@@

The most striking feature of the mass ratio distribution shown in Figure \ref{fig:mrd} is the turnover or flattening at intermediate q. The maximum of the lognormal distribution occurs at $q = 0.30 \pm 0.03$ and is an estimate of the characteristic scale.

@@

The integrals are estimated as part of the nested sampling algorithm in the MultiNest code. The odds ratio comparing the lognormal distribution to power law fit is $Z_\mathrm{lognormal} / Z_\mathrm{power} = 5.1 \pm 0.1 \e{6}$, indicating a very strong preference for the lognormal distribution model. We also compare to the mass ratio distribution expected for random pairing from the Kroupa IMF and to a uniform distribution (a special case of the power law). In these cases, there are no free parameters so the evidence integral just becomes the likelihood function (Equation \ref{eqn:money}). The corresponding odds ratios are  $Z_\mathrm{lognormal} / Z_\mathrm{IMF} = 6.5 \pm 0.1 \e{22}$ and $Z_\mathrm{lognormal} / Z_\mathrm{uniform} = 7.0 \pm 0.1 \e{6}$. Both of these again demonstrate a very strong preference of the lognormal distribution.

@@

Our mass ratio distribution appears to be in tension with the results of the VAST survey \citep{DeRosa2014}, which finds a nearly flat distribution for close companions. However, their subsample of close companions only includes 18 binaries, so it is possible that the different forms are just a result of small number statistics. To assess the degree of tension, we use the Anderson-Darling  test \citep{Anderson1954} to find the probability that both their close companion subsample and our companions are drawn from the same parent distribution. We only use companions from this work with mean $\overline{q} > 0.15$ because the VAST survey subsample makes the same cut. To account for measurement uncertainties in the mass ratios, we choose from the mass ratio samples ($q_k^{(n)}$, see Section \ref{sec:mrd}) many times and compute the Anderson-Darling test statistic each time. Since \citet{DeRosa2014} do not quote uncertainties, we assume uncertainties of $\sigma_q = 0.05$ for all of their measurements. The result is $p = 0.10^{+0.07}_{-0.04}$; we cannot reject the hypothesis that both samples come from the same distribution.

@@

The mass ratio distribution derived in this work has a very different form than the power law found for companions at wide separations. This is likely a result of disk interactions as the two components are accreting. The close companions that we detect may form with similar masses to their counterparts at large separations ($a \gtrsim 1000\ AU$), but preferentially accrete matter from the dense primary star disk. Since preferential accretion is strongest at the most extreme mass ratios [CITE], the result would be a depletion of low mass ratio companions as they become intermediate to high mass ratio companions. The characteristic scale of $\sim 0.3$ that we see in Figure \ref{fig:mrd} would then be related to the disk timescale, since with enough time the preferential accretion would push all companions to $q = 1.0$.

It is also possible that some of the companions found in this work were formed from a gravitationally unstable disk \citep[e.g.][]{Kratter2006, Stamatellos2011}. Being a completely different formation mechanism than the way wide companions form, we would expect the mass function to differ. Additionally, such companions would undergo the same preferential accretion discussed above. This explanation may even be preferable, since it could explain why we see a separation-dependent mass ratio distribution around intermediate-mass stars but not low-mass stars: more massive stars tend to have more massive disks [CITE] that are more likely to fragment.

@@

In this work, we described a binary survey of 341 bright A- and B-type stars. We used the direct spectroscopic detection method \citep{Gullikson2016} to find the spectral lines of 64 companions with temperatures ranging from $3600 - 16000\ K$.  We used the cross-correlation functions to estimate the temperature, surface gravity, and metallicity of most of our sample stars, and converted to mass and age by interpolating Padova stellar evolutionary tracks \citep{Bressan2012}. Likewise, we convert the companion temperature measurements to mass by using solar metallicity Dartmough evolutionary tracks [CITE].

We then use the formalism introduced in \citet{Foreman2014}, which self-consistently accounts for measurement errors, to infer the form of the mass ratio distribution (shown in Figure \ref{fig:mrd}). Unlike most previous work, we find that a power law is a poor descriptor of the data and find that a lognormal distribution performs much better. This result, which only includes close companions since it is a spectroscopic technique, is consistent with the 18 close companions found in the VAST survey \cite{DeRosa2014}. However, this result shows much more detail due to a larger number of companions.

We interpret the mass ratio distribution in terms of formation mechanism in Section \ref{subsec:theory}. It is likely that the mass ratio distribution we find is largely a result of preferential accretion onto the secondary star, which largely stops when the circumprimary or circumbinary disk dissipates.

>>>>

\begin{abstract}
Binary stars and higher-order multiple systems are a ubiquitous outcome of star formation, especially as the system mass increases. The distribution of mass ratios is a unique probe into the conditions of the collapsing cloud core and circumstellar disk(s) of the binary fragments. Inside $a \sim 1000$ AU the disks from the two forming stars can interact, and additionally companions can form directly through disk fragmentation. We should therefore expect the mass-ratio distribution of close companions to differ from that of wide companions. This prediction is difficult to test with intermediate-mass to massive primary stars using traditional methods because close companions are so much dimmer. We present the results of a survey searching for companions to A- and B-type stars using the direct spectral detection technique, which is sensitive to late-type companions within $\sim 1''$ of the primary and which has no inner working angle. We estimate the temperatures and surface gravity of most of the 341 sample stars, and derive their masses and ages. We additionally estimate the temperatures and masses of the 64 companions we find, 23 of which are new detections. We find that the mass-ratio distribution for our sample has a maximum near $q \sim 0.3$. Our mass-ratio distribution has a very different form than in previous work, where it is usually well-described by a power law, and indicates that close companions to intermediate-mass stars experience significantly different accretion histories or formation mechanisms than wide companions.
\end{abstract}
\keywords{binaries: spectroscopic, stars:early-type, stars:formation, stars:statistics}

@@

The overall multiplicity rate and the distributions of mass ratio, period, and eccentricity of a binary star population place important constraints on the mode of binary star formation. While the period and eccentricity are altered by dynamical processing in the birth cluster, the present-day mass ratio of a binary system is a direct result of its formation \citep{Parker2013}. Most binary stars are thought to form via core fragmentation \citep{Boss1979, Boss1986, Bate1995}, in which a collapsing core fragments into two or more individual protostars. The number and initial masses of the fragments are set by the total core mass, as well as its rotation, turbulence, and its temperature and density structure. If the fragments are well separated ($a \gtrsim 1000$ AU), they will evolve independently of each other, accreting mass from the core material onto their own protostellar disks and then onto the protostars themselves. However close fragments ($a \sim 100$ AU) will interact with each other; the protostellar disk may be truncated, destabilized, or form into a circumbinary disk if the separation is small enough \citep{Bate1997}. In addition, an unstable disk can fragment to form low-mass companions \citep{Kratter2006, Stamatellos2011}. The mass ratios of close companions formed via either mechanism should be affected by preferential accretion. Most work has suggested that the disk material will preferentially accrete onto the lower mass companion \citep{Bate1997, BBB2002}; however, recent work has indicated that magnetic disk braking may result in preferential accretion onto the more massive component \citep{Zhao2013} instead. In either case, we would expect to find a mass-ratio distribution for companions inside a few $100$ AU that differs from that of companions on wider orbits.

The mass ratio, period, and eccentricity distributions are fairly well-known for solar type stars \citep{Duquennoy1991, Raghavan2010} and cooler stars \citep{Fischer1992, Delfosse2004}. Interestingly, the mass-ratio distribution appears to be invariant to separation for these stars \citep{Meyer2013}, contrary to the theoretical expectations. All of the distributions are much less certain for more massive stars. The reason for this is two-fold: first, more massive stars tend to more rare and farther away, meaning many of the companions are angularly close to the very bright primaries and difficult to detect with imaging techniques. Second, the primary stars tend to be rapid rotators, which limits radial velocity precision to $\sim 1 \mathrm{km\ s}^{-1}$. Additionally, radial velocity monitoring can only measure a mass ratio if spectral lines from both components are visible; this typically suffers from the same flux ratio difficulty as imaging techniques.

Nonetheless, \citet{DeRosa2014} performed an adaptive optics imaging survey of nearby A-type stars, and found that the mass-ratio distribution is well-described by a power law with large slope, indicating a very strong preference for low-mass companions. They also found initial evidence that the mass-ratio distribution for companions inside $125$ AU has a much shallower power law slope than that of wide companions, and is consistent with flat. Their close companion subsample contained only 18 binary systems, and the result is complicated by the inherent difficulty of detecting close companions with low mass ratios in an imaging survey.

@@

In this paper, we describe a spectroscopic survey of nearby chemically normal, main sequence intermediate-mass stars ($M \approx 1.5 - 15 M_{\odot}$). We search for companions using the direct spectral detection technique \citep{Gullikson2016}, which has a separation-invariant detection rate for all separations inside $\sim 1 ''$. We describe the stellar sample and data used for the survey in Section \ref{sec:obs}, as well as the data reduction steps in the same section. Next, we describe the direct spectral detection method and tabulate the companion detections in Section \ref{sec:companions}. We estimate the mass and age of the primary stars in Section \ref{sec:sp}, and discuss the survey completeness in Section \ref{sec:completeness}. Finally, we end with a derivation of the mass-ratio distribution from our sample in Section \ref{sec:mrd} and discuss its implications for binary formation in Section \ref{sec:discussion}.

@@

We use the methodology developed in \citet{Foreman2014} to perform bayesian inference on the shape and form of the mass-ratio distribution. The log-likelihood function in this formalism is derived from modeling the survey as a draw from the inhomogeneous Poisson process with rate density $\Gamma \equiv KQ(q)P(q)$:

@@

\begin{align}
P(\vec{\theta}| \alpha, m, \tau, \epsilon) &= \mathcal{N}(\vec{\theta} | m, K(\{\Delta_j\}, \alpha, \tau, \epsilon)) \\
K_{ij} &= \sqrt{\left[\alpha \exp{\left(-\frac{(\Delta_i - \Delta_j)^2}{2\tau^2}\right)}\right]^2 + \epsilon^2 \delta_{ij}}
\end{align}
The smoothing prior is an 7-dimensional gaussian with mean $m$ and covariance matrix $K_{ij}$, and encodes our belief that the mass-ratio distribution is a smoothly varying function while leaving enough flexibility to let the data drive the shape of the function. Since we have introduced three new hyperparameters ($a, m, \tau, \epsilon$), we must apply a prior to them and marginalize over them when estimating the bin heights. We choose log-uniform priors for $a, \tau$, and $\epsilon$, and a uniform prior for the mean $m$. The full posterior probability distribution for the histogram model is:

@@

We show the resulting fits in Figure \ref{fig:mrd}. The $1\sigma$ uncertainties in the bin heights from the histogram model are shown as error bars, and we overplot 300 samples of the lognormal distribution fit to show the spread allowed by the data. The best-fit power law is plotted with a red dot-dashed line. We also estimate the mass-ratio distribution expected from random pairing of the Kroupa Initial Mass Function (IMF), and show the result in yellow. We estimated the distribution by drawing 100000 primary stars from the IMF with masses between $1.5 < M/M_{\odot} < 20$. We then drew companions from the same IMF, with the restriction that the companion has a lower mass than the primary. The result plotted in yellow in Figure \ref{fig:mrd} is a gaussian kernel density estimate of the resulting mass ratios, with a bandwidth of $0.05$.

\begin{figure}
\includegraphics[width=\columnwidth]{MRD_total.pdf}
\caption{mass-ratio distribution for our sample. The data was fit to a histogram, a lognormal distribution, and a power law. The histogram is shown in the solid blue blocks, with $1 \sigma$ uncertainties marked with error bars. The variance of the lognormal  fit is shown with 300 samples from the posterior probability distribution for the parameters in green. We also show the best-fit power law and the mass-ratio distribution resulting from random pairing of the Kroupa initial mass function. Finally, we display the raw mass ratio measurements with associated uncertainties in the cluster of data points near the top of the figure.}
\label{fig:mrd}
\end{figure}

@@

The most striking feature of the mass-ratio distribution shown in Figure \ref{fig:mrd} is the turnover or flattening at intermediate q. The maximum of the lognormal distribution occurs at $q = 0.30 \pm 0.03$ and is an estimate of the characteristic scale.

@@

The integrals are estimated as part of the nested sampling algorithm in the MultiNest code. The odds ratio comparing the lognormal distribution to power law fit is $Z_\mathrm{lognormal} / Z_\mathrm{power} = 5.1 \pm 0.1 \e{6}$, indicating a very strong preference for the lognormal distribution model. We also compare to the mass-ratio distribution expected for random pairing from the Kroupa IMF and to a uniform distribution (a special case of the power law). In these cases, there are no free parameters so the evidence integral just becomes the likelihood function (Equation \ref{eqn:money}). The corresponding odds ratios are  $Z_\mathrm{lognormal} / Z_\mathrm{IMF} = 6.5 \pm 0.1 \e{22}$ and $Z_\mathrm{lognormal} / Z_\mathrm{uniform} = 7.0 \pm 0.1 \e{6}$. Both of these again demonstrate a very strong preference of the lognormal distribution.

@@

Our mass-ratio distribution appears to be in tension with the results of the VAST survey \citep{DeRosa2014}, which finds a nearly flat distribution for close companions. However, their subsample of close companions only includes 18 binaries, so it is possible that the different forms are just a result of small number statistics. To assess the degree of tension, we use the Anderson-Darling  test \citep{Anderson1954} to find the probability that both their close companion subsample and our companions are drawn from the same parent distribution. We only use companions from this work with mean $\overline{q} > 0.15$ because the VAST survey subsample makes the same cut. To account for measurement uncertainties in the mass ratios, we choose from the mass ratio samples ($q_k^{(n)}$, see Section \ref{sec:mrd}) many times and compute the Anderson-Darling test statistic each time. Since \citet{DeRosa2014} do not quote uncertainties, we assume uncertainties of $\sigma_q = 0.05$ for all of their measurements. The result is $p = 0.10^{+0.07}_{-0.04}$; we cannot reject the hypothesis that both samples come from the same distribution.

@@

The mass-ratio distribution derived in this work has a very different form than the power law found for companions at wide separations. This is likely a result of disk interactions as the two components are accreting. The close companions that we detect may form with similar masses to their counterparts at large separations ($a \gtrsim 1000\ AU$), but preferentially accrete matter from the dense primary star disk. Since preferential accretion is strongest at the most extreme mass ratios [CITE], the result would be a depletion of low mass ratio companions as they become intermediate to high mass ratio companions. The characteristic scale of $\sim 0.3$ that we see in Figure \ref{fig:mrd} would then be related to the disk timescale, since with enough time the preferential accretion would push all companions to $q = 1.0$.

It is also possible that some of the companions found in this work were formed from a gravitationally unstable disk \citep[e.g.][]{Kratter2006, Stamatellos2011}. Being a completely different formation mechanism than the way wide companions form, we would expect the mass function to differ. Additionally, such companions would undergo the same preferential accretion discussed above. This explanation may even be preferable, since it could explain why we see a separation-dependent mass-ratio distribution around intermediate-mass stars but not low-mass stars: more massive stars tend to have more massive disks [CITE] that are more likely to fragment.

@@

In this work, we described a binary survey of 341 bright A- and B-type stars. We used the direct spectral detection method \citep{Gullikson2016} to find the spectral lines of 64 companions with temperatures ranging from $3600 - 16000\ K$.  We used the cross-correlation functions to estimate the temperature, surface gravity, and metallicity of most of our sample stars, and converted to mass and age by interpolating Padova stellar evolutionary tracks \citep{Bressan2012}. Likewise, we convert the companion temperature measurements to mass by using solar metallicity Dartmough evolutionary tracks [CITE].

We then use the formalism introduced in \citet{Foreman2014}, which self-consistently accounts for measurement errors, to infer the form of the mass-ratio distribution (shown in Figure \ref{fig:mrd}). Unlike most previous work, we find that a power law is a poor descriptor of the data and find that a lognormal distribution performs much better. This result, which only includes close companions since it is a spectroscopic technique, is consistent with the 18 close companions found in the VAST survey \cite{DeRosa2014}. However, this result shows much more detail due to a larger number of companions.

We interpret the mass-ratio distribution in terms of formation mechanism in Section \ref{subsec:theory}. It is likely that the mass-ratio distribution we find is largely a result of preferential accretion onto the secondary star, which largely stops when the circumprimary or circumbinary disk dissipates.
